### 2024-04-18 manuscript.qmd
first commit

---
title: "Machine Learning Approaches for Ammonia Volatilization Prediction after Manure Application"
author: 
  - "Armand Favrot"
  - "Sophie Génermont"
  - "Céline Decuq"
  - "Jean-Marc Giliot"
  - "David Makowski"
format:
  html:
    toc: true
    embed-resources: true
---

```{r setup, include=FALSE}
# Options pour le rmarkdown
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r}
library ("ggplot2") 
library ("gridExtra")
library ("dplyr")       
library ("paletteer")       
library ("tibble") # rownames_to_column
library ("tidyr") # pivot_longer
library ("ALFAM2")
library ("randomForest")
library ("xgboost")
library ("glmnet")
library ("shapviz")
library ("treeshap")

set.seed (123)
```

```{r}
axis_text_size = 24 
axis_title_size = 30
title_size = 26

theme_replace(
    axis.text = element_text(size = axis_text_size), 
    axis.title.x = element_text(size = axis_title_size, angle = 0, margin = ggplot2::margin (t = 10)), 
    axis.title.y = element_text(size = axis_title_size, angle = 90, margin = ggplot2::margin (r = 10)), 
    plot.title = element_text(size = title_size, hjust = 0.5, face = "bold"), 
    legend.text = element_text(size = axis_title_size), 
    legend.title = element_text(size = axis_title_size, face = "bold"), 
    strip.text = element_text(size = axis_title_size, face = "bold", color = "white"), 
    strip.background = element_rect(fill = "#000080"), 
    panel.background = element_rect(fill = "#f3faff"), panel.grid.major = element_line(colour = "white"), 
    panel.spacing = unit(2, "lines")
)

options(ggplot2.discrete.fill = list("#67c5a7"))

Dark2 = paletteer_d("RColorBrewer::Dark2")
```

# Data description

```{r}
load (file = "scripts/data_alfam2.Rdata")
```

```{r}
plot = data_alfam2 %>%

    filter (! (app.mthd == "bsth" & incorp == "shallow")) %>%

    mutate (strategy = paste (app.mthd, incorp, sep = " - ")) %>%
    mutate (strategy = recode (strategy, "bc - none" = "A", "bc - shallow" = "B", "bc - deep" = "C", "bsth - none" = "D", "os - none" = "E", "ts - none" = "F")) %>%

    ggplot () +
        geom_line (aes (x = time, y = e.rel, group = pmid), linewidth = 0.15) +
        ylab ("Cumul of relative emission") +
        xlab ("Time (h)") +
        theme (axis.title.y = element_text (margin = ggplot2::margin (r = 25))) +
        # scale_color_manual (values = Dark2) +
        facet_wrap (~ strategy)

png (file = "figures/data_description_relative_cumul.png", width = 1000, height = 700)
plot
dev.off()

plot
```

# Part 1 - Model comparison

## Alfam2

```{r}
pars <- alfam2pars01[!grepl('man.ph', names(alfam2pars01))]

alfam2_predictions =  alfam2 (

        # Important to use alfam2pars01 parameters, which correspond to the parameters used in the paper
        # New parameters are available, but leads to ohter results (less good results)
        pars = pars,

        dat = data_alfam2 %>% select (- j.NH3, - e.cum, - e.rel, - dataset),

        app.name = "tan.app",

        time.name = "time",

        time.incorp = "t.incorp",

        group = "pmid",

        prep = TRUE,

        warn = FALSE
)
```

We add the true values to the prediction dataframe and we keep only the last observation of each pmid.

```{r}
alfam2_predictions = alfam2_predictions %>%

        select (pmid, time, e, er) %>%

        mutate (truth_e = data_alfam2$e.cum) %>%
        mutate (truth_er = data_alfam2$e.rel) %>%

        mutate (dataset = data_alfam2$dataset) %>%

        select (pmid, time, e, er, truth_e, truth_er, dataset) %>%

        filter (time == max (time), .by = pmid)
```

## Random forest

```{r}
load (file = "scripts/data_random_forest.Rdata")
```

```{r}
data_random_forest_calibration = data_random_forest %>% 
    filter (dataset == "Calibration subset") %>%
    select (- pmid, - e.rel, - dataset)

random_forest_model = randomForest (

        e.cum ~ .,

        data = data_random_forest_calibration,

        importance = TRUE,

        mtry = 22, nodesize = 3, replace = FALSE, sample_frac = 0.8

)
```

```{r}
plot (random_forest_model)
```

```{r}
random_forest_predictions = data_random_forest %>%

    mutate (e.cum_hat = predict ( 
                random_forest_model,                       
                newdata = data_random_forest %>% 
                    select (- pmid, - e.cum, - e.rel, - dataset)
            )
    ) %>%

    mutate (e.rel_hat = e.cum_hat / tan.app) %>%

    select (pmid, time, e = e.cum_hat, er = e.rel_hat, truth_e = e.cum, truth_er = e.rel, dataset)    
```

## Xgboost

```{r}
load (file = "scripts/data_xgboost.Rdata")
```

```{r}

xgboost_model = xgboost (

    data = xgb.DMatrix (

        data = data_xgboost %>% 
            filter (dataset == "Calibration subset") %>%
            select (- pmid, - e.cum, - e.rel, - dataset) %>%
            as.matrix %>%
            {.},

        label = data_xgboost %>% 
            filter (dataset == "Calibration subset") %>%
            select (e.cum) %>% 
            as.matrix %>%
            {.}

    ),

    max.depth = 4, nrounds = 150, eta = 0.3, min_child_weight = 0.3, subsample = 1,

    verbose = FALSE,

    objective = "reg:squarederror"

)
```

```{r}
xgboost_predictions = data_xgboost %>%

    mutate (e.cum_hat = predict (

                xgboost_model, 
                newdata = data_xgboost %>% 
                        select (- pmid, - e.cum, - e.rel, - dataset) %>% 
                        as.matrix
            )
    ) %>%

    mutate (e.rel_hat = e.cum_hat / tan.app) %>%

    select (pmid, time, e = e.cum_hat, er = e.rel_hat, truth_e = e.cum, truth_er = e.rel, dataset)
```

## Lasso

```{r}
load (file = "scripts/data_lasso.Rdata")
```

```{r}
lasso_model = cv.glmnet (

    x = data_lasso %>% 
        filter (dataset == "Calibration subset") %>%
        select (- pmid, - e.cum, - e.rel, - dataset) %>%
        as.matrix %>%
        {.},

    y = data_lasso %>% 
        filter (dataset == "Calibration subset") %>%
        select (e.cum) %>% 
        as.matrix %>%
        {.},

    alpha = 1
)
```

```{r}
lasso_predictions_vector = predict (

    lasso_model,

    data_lasso %>% 
            select (- pmid, - e.cum, - e.rel, - dataset) %>% 
            as.matrix
)

lasso_predictions = data_lasso %>%

    mutate (e.cum_hat = exp (lasso_predictions_vector), e.cum = exp (e.cum)) %>%

    mutate (e.rel_hat = e.cum_hat / tan.app) %>%

    select (pmid, time, e = e.cum_hat, er = e.rel_hat, truth_e = e.cum, truth_er = e.rel, dataset) 
```

## Comparison

```{r}
predictions_of_all_methods = rbind (
    alfam2_predictions %>% mutate (method = "alfam2"), 
    random_forest_predictions %>% mutate (method = "random forest"),
    xgboost_predictions %>% mutate (method = "xgboost"),
    lasso_predictions %>% mutate (method = "lasso")
)
```

### Observed vs predicted values

```{r}
plot = predictions_of_all_methods %>%

    filter (dataset == "Evaluation subset") %>%

    mutate (method = recode (method, "alfam2" = "A", "lasso" = "B", "random forest" = "C", "xgboost" = "D")) %>%

    ggplot () +

        geom_point (aes (x = truth_e, y = e, color = method), size = 3) +

        geom_abline (slope = 1, linetype = "dashed") +

        facet_wrap (~ method) +

        scale_color_manual (values = Dark2[c(2, 5, 6, 8)]) +

        labs (color = "") +

        theme (legend.position = "none",
               axis.title.y = element_text (margin = ggplot2::margin (r = 30)),
               axis.title.x = element_text (margin = ggplot2::margin (t = 15, b = 20))) +

        xlab ("Observed values") + ylab ("Predicted values") +

        NULL

png (file = "figures/observed_vs_predicted_values.png", width = 1000, height = 700)
plot
dev.off()

plot
```

## Evaluation metrics

```{r}
df_evaluation_metrics = rbind (

    predictions_of_all_methods %>% 
        select (prediction = e, truth = truth_e, dataset, method) %>% 
        mutate (name = "72h cum. emission"),

    predictions_of_all_methods %>% 
        select (prediction = er, truth = truth_er, dataset, method) %>% 
        mutate (name = "72h relative cum. emission")

) %>%

    summarise (

        Pearsons_r = cor (prediction, truth),

        ME = 1 - (sum ( (prediction - truth) ^ 2) / sum ( (truth - mean (truth)) ^ 2)),

        MAE = mean (abs (prediction - truth)),

        MBE = mean (prediction - truth),

        .by = c (name, dataset, method)

) %>%

    mutate_if (is.numeric, round, digits = 3)

df_evaluation_metrics
```

```{r}
df_evaluation_metrics %>%

    rename (response = name) %>%
    pivot_longer (cols = c (Pearsons_r, ME, MAE, MBE)) %>%
    mutate (name = factor (name, levels = c("Pearsons_r", "ME", "MAE", "MBE"))) %>% 
    filter (name %in% c ("MAE", "MBE")) %>%

    arrange (name, dataset) %>%

    mutate (variation = (1 - abs(value) / abs(value [method == "alfam2"])) * 100, .by = c(response, name, dataset))
```

```{r}
# Evaluation for 72h cum. emission --------
df_plot = df_evaluation_metrics %>%
    filter (name == "72h cum. emission" & dataset == "Evaluation subset") %>%
    select (- name) %>%
    pivot_longer (cols = c (Pearsons_r, ME, MAE, MBE)) %>%
    mutate (name = factor (name, levels = c("Pearsons_r", "ME", "MAE", "MBE")))

suppressWarnings(ggplot (df_plot) +
    geom_histogram (aes (x = name, y = value, fill = method), position = "dodge", stat = "identity") +
    facet_wrap (~ name, scales = "free", nrow = 1) + 
    xlab ("") + ylab ("") + labs (fill = "") +
    theme (legend.position = "none ", strip.text.y = element_text(angle = 0), axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
    scale_fill_manual (values = Dark2[c(2, 5, 6, 8)]) +
    ggtitle ("Evaluation subset") + theme (plot.title = element_text (margin= ggplot2::margin(0,0,20,0), hjust = 0, size = 30)))
# --------

# Evaluation for 72h relative cum. emission ----------
df_plot = df_evaluation_metrics %>%
    filter (name == "72h cum. emission" & dataset == "Calibration subset") %>%
    select (- name) %>%
    pivot_longer (cols = c (Pearsons_r, ME, MAE, MBE)) %>%
    mutate (name = factor (name, levels = c("Pearsons_r", "ME", "MAE", "MBE")))

suppressWarnings(ggplot (df_plot) +
    geom_histogram (aes (x = name, y = value, fill = method), position = "dodge", stat = "identity") +
    facet_wrap (~ name, scales = "free", nrow = 1) + 
    xlab ("") + ylab ("") + labs (fill = "") +
    theme (legend.position = "bottom", strip.text.y = element_text(angle = 0), axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
    scale_fill_manual (values = Dark2[c(2, 5, 6, 8)]) +
    ggtitle ("Calibration subset") + theme (plot.title = element_text (margin= ggplot2::margin(0,0,20,0), hjust = 0, size = 30)))
# --------
```

# Part 2 - New training with the whole dataset

## Random forest

```{r}
random_forest_model = randomForest (

        e.cum ~ .,

        data = data_random_forest %>% select (- pmid, - e.rel, - dataset),

        importance = TRUE,

        mtry = 22, nodesize = 3, replace = FALSE, sample_frac = 0.8

)
```

### Shapley

```{r}
unified_model_rf <- randomForest.unify(random_forest_model, data_random_forest_calibration %>% select (- e.cum))

treeshap_rf <- treeshap(unified_model_rf, data_random_forest_calibration %>% select (- e.cum))

shapley_rf <- shapviz(treeshap_rf, X = data_random_forest_calibration %>% select (- e.cum))
```

```{r}
plot_shap_rf = sv_importance(shapley_rf, kind = "beeswarm", max_display = 10L) +
    scale_colour_gradient(low = "darkblue", high = "red", breaks = c(0, 1), labels = c("low", "high")) +
    theme (
      legend.title = element_text (size = 25, margin = ggplot2::margin (r = 50, b = 100)),
      legend.position = "bottom",
      legend.key.width = unit(3, "cm"),
      axis.title.x = element_text (margin = ggplot2::margin( b = 30))
    )

plot_shap_rf
```

### Gain

```{r}
df_importance_rf = importance (random_forest_model) %>%
    as.data.frame %>%
    rownames_to_column (var = "Variable") %>%
    mutate (`%IncMSE` = `%IncMSE` / max (`%IncMSE`)) %>%
    mutate (type = "A")

plot_importance_rf = ggplot (df_importance_rf) + 
        geom_point (aes (x = `%IncMSE`, y = Variable), size = 5) +
        scale_y_discrete (limits = df_importance_rf %>% arrange (`%IncMSE`) %>% pull (Variable)) +
        xlab ("Importance") +
        ylab ("") +
        facet_wrap (~ type)

plot_importance_rf
```
## Xgboost

### Shapley

```{r}
xgboost_model = xgboost (

    data = xgb.DMatrix (

        data = data_xgboost %>% 
            select (- pmid, - e.cum, - e.rel, - dataset) %>%
            as.matrix %>%
            {.},

        label = data_xgboost %>% 
            select (e.cum) %>% 
            as.matrix %>%
            {.}

    ),

    max.depth = 4, nrounds = 150, eta = 0.3, min_child_weight = 0.3, subsample = 1,

    verbose = FALSE,

    objective = "reg:squarederror"

)
```

```{r}
df_tmp = data_xgboost %>% select (- pmid, - e.cum, - e.rel, - dataset)

unified_model_xgboost <- xgboost.unify(xgboost_model, as.matrix(df_tmp))

treeshap_xgboost <- treeshap(unified_model_xgboost, df_tmp)

shapley_xgb <- shapviz(treeshap_xgboost, X = df_tmp)
```

```{r}
plot_shap_xgboost = sv_importance(shapley_xgb, kind = "beeswarm", max_display = 10L) +
  scale_colour_gradient(low = "darkblue", high = "red", breaks = c(0, 1), labels = c("low", "high")) +
  theme (legend.title = element_text (size = 30, margin = ggplot2::margin (r = 50, b = 100)),
         legend.position = "bottom",
         legend.key.width = unit(3, "cm"),
         axis.title.x = element_text (margin = ggplot2::margin( b = 30)))

plot_shap_xgboost
```

### Gain

```{r}
df_importance_xgb = xgb.importance (model = xgboost_model) %>%
    mutate (Gain = Gain / max (Gain)) %>%
    mutate (type = "B")

plot_importance_xgboost = df_importance_xgb %>%

    ggplot() +
        geom_point (aes (x = Gain, y = Feature), size = 5) +
        scale_y_discrete (limits = df_importance_xgb %>% arrange (Gain) %>% pull (Feature)) +
        xlab ("Importance") + ylab ("") +
        facet_wrap (~ type)

plot_importance_xgboost
```
# Part 3 - Testing scenarios

```{r}
load (file = "scripts/data_scenarios.Rdata")

data_scenarios = data_scenarios %>%

    mutate (t.incorp = as.factor (t.incorp)) %>%
    mutate (t.incorp = recode (t.incorp, "1000" = "-"))
```

## Random forest

```{r}
load (file = "scripts/data_scenarios_random_forest.Rdata")

data_tmp = data_scenarios_random_forest

random_forest_predictions_scenarios_vector = predict (
        random_forest_model,       
        newdata = rbind (data_random_forest_calibration[1, -1], data_tmp)[- 1, ]
)

data_scenarios = data_scenarios %>%

    mutate (e.cum_hat_random_forest = random_forest_predictions_scenarios_vector, .before = time) %>%

    mutate (efficacy_random_forest = ((e.cum_hat_random_forest / e.cum_hat_random_forest[app.mthd == "bc" & incorp == "none" & t.incorp == "-"]) - 1) * 100, 
            .by = c (tan.app, app.rate, man.dm, man.ph, man.source, group_temp, group_wind, group_rain)) %>%

    {.}
```

## Xgboost

```{r}
load (file = "scripts/data_scenarios_xgboost.Rdata")

xgboost_predictions_scenarios_vector = predict (

    xgboost_model,

    data_scenarios_xgboost %>% as.matrix
)

data_scenarios = data_scenarios %>%

    mutate (e.cum_hat_xgboost = xgboost_predictions_scenarios_vector, .before = time) %>%
    mutate (e.cum_hat_xgboost = ifelse (e.cum_hat_xgboost < 0, 0, e.cum_hat_xgboost)) %>%

    mutate (efficacy_xgboost = ((e.cum_hat_xgboost / e.cum_hat_xgboost[app.mthd == "bc" & incorp == "none" & t.incorp == "-"]) - 1) * 100, 
            .by = c (tan.app, app.rate, man.dm, man.ph, man.source, group_temp, group_wind, group_rain))
```

## Lasso

```{r}
load (file = "scripts/data_scenarios_lasso.Rdata")

lasso_predictions_scenarios_vector = predict (

    lasso_model,

    data_scenarios_lasso %>% as.matrix
)

data_scenarios = data_scenarios %>%

    mutate (e.cum_hat_lasso = lasso_predictions_scenarios_vector, .before = time) %>%

    mutate (efficacy_lasso = ((e.cum_hat_lasso / e.cum_hat_lasso[app.mthd == "bc" & incorp == "none"]) - 1) * 100, 
            .by = c (tan.app, app.rate, man.dm, man.ph, man.source, group_temp, group_wind, group_rain))
```

## Alfam2

```{r}
load (file = "scripts/data_scenarios_alfam2.Rdata")

alfam2_predictions_scenarios_vector =  alfam2 (
        pars = alfam2pars01,
        dat = data_scenarios_alfam2,
        app.name = "tan.app",
        time.name = "time",
        time.incorp = "t.incorp",
        group = "pmid",
        prep = TRUE,
        warn = FALSE
)  %>%

    filter (time == max (time), .by = pmid) %>%
    select (pmid, e.cum_hat_alfam2 = e)

data_scenarios = data_scenarios %>%

    left_join (alfam2_predictions_scenarios_vector, by = "pmid") %>%

    mutate (efficacy_alfam2 = ((e.cum_hat_alfam2 / e.cum_hat_alfam2[app.mthd == "bc" & incorp == "none" & t.incorp == "-"]) - 1) * 100, 
            .by = c (tan.app, app.rate, man.dm, man.ph, man.source, group_temp, group_wind, group_rain))
```

## Comparison

```{r}
plot = data_scenarios %>% 
    select (app.mthd, incorp, t.incorp, man.ph,
            efficacy_random_forest, efficacy_xgboost, efficacy_lasso, efficacy_alfam2) %>%

    rename (efficacy_rforest = efficacy_random_forest) %>%

    pivot_longer (cols = c (5 : 8), names_to = c (".value", "method"), names_pattern = "(.+)_(.+)") %>%
    mutate (method = recode (method, "rforest" = "random forest")) %>%

    mutate (method = recode (method, "alfam2" = "A", "lasso" = "B", "random forest" = "C", "xgboost" = "D")) %>%

    mutate (group = paste (app.mthd, incorp, t.incorp)) %>%
    filter (! (app.mthd == "bc" & incorp == "none")) %>%

    mutate (group = recode (group,
                            "bc shallow 0" = "incorporation",
                            "bsth none -" = "trailing hose",
                            "os none -" = "open slot",
                            "ts none -" = "trailing shoe")) %>%

    ggplot () +

        geom_boxplot (aes (x = efficacy, y = group, fill = group)) +
        scale_fill_manual (values = Dark2[c(2 : 5)]) +
        ylab ("") + xlab ("Efficacy") +
        theme (legend.position = "none",
               axis.ticks.y = element_blank()) +
        facet_wrap (~ method, ncol = 1)

png (file = "figures/emissions_for_scenarios.png", width = 1000, height = 800)
plot
dev.off()

plot
```

# Figures

```{r}
legend = cowplot::get_legend (plot_shap_rf)

plot_shap_rf_2 = plot_shap_rf +
    annotate (geom = "label", label = "A", size = 22, fontface = "bold", x = 58, y = 1.5, fill = "black", color = "white") +
    xlim (c (- 20, 58)) +
    theme (legend.position = "none", 
          plot.margin = unit (c (0, 0, 1, 0), "cm"),
          axis.text.x = element_blank(),
          axis.title.x = element_blank(),
          axis.ticks.x = element_blank())

plot_shap_xgboost_2 = plot_shap_xgboost +
    annotate (geom = "label", label = "B", size = 22, fontface = "bold", x = 58, y = 1.5, fill = "black", color = "white") +
    xlim (c (- 20, 58)) + 
    theme (legend.position = "none", 
          plot.margin = unit (c (0, 0, 1, 0), "cm"),
          axis.title.x = element_text (margin = ggplot2::margin (t = 10, b = 20)))

png (file = "figures/plot_shap.png", width = 1000, height = 900)
do.call ("grid.arrange", 
         c (list(plot_shap_rf_2, 
                 plot_shap_xgboost_2, 
                 legend), 
            list(layout_matrix = as.matrix(c (1, 2, 3)),
                  heights = c (0.8, 1, 0.2))))
dev.off()
```

```{r}
png (file = "figures/plot_importance.png", width = 1300, height = 800)
grid.arrange (plot_importance_rf, plot_importance_xgboost, nrow = 1)
dev.off()
```

# Supplementary material