### 2024-04-20 scripts/random_forest_cross_validation.qmd

---
title: "Random forest cross-validation"
format:
  html:
    toc: true
    embed-resources: true
---

```{r setup, include=FALSE}
# Options pour le rmarkdown
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r}
library (dplyr)
library (randomForest)
```

```{r}
load (file = "processed_data/data_random_forest.Rdata")
```

```{r}
data_random_forest_calibration = data_random_forest %>% 
    filter (dataset == "Calibration subset") %>%
    select (- pmid, - e.rel, - dataset)
```

```{r}
# set.seed (123)

# for (i in c (1 : 490)){
# 
#     calibration_subset_train = data_random_forest_calibration[- i, ]
# 
#     calibration_subset_eval = data_random_forest_calibration[i, ]
# 
#     # create hyper_grid search ------
#     hyper_grid <- expand.grid (
# 
#         mtry = c (8 : 28),
#         nodesize = c(1, 3, 5),
#         replace = c(TRUE, FALSE),
#         samp.fraction = c(0.63, 0.8, 1),
#         prediction = NA,
#         truth = NA,
#         executation_time = NA,
#         row_index = i
#     ) %>%
# 
#         filter (! (samp.fraction == 1 & replace == FALSE)) %>%
# 
#         mutate (sampsize = floor (samp.fraction * nrow (calibration_subset_train)))
#     # ------
# 
#     for (j in seq_len(nrow(hyper_grid))) {
# 
#         t0 = Sys.time()
# 
#         # fit model for jth hyperparameter combination
#         fit <- randomForest(
#                 formula      = e.cum ~ ., 
#                 data         = calibration_subset_train, 
#                 mtry         = hyper_grid$mtry[j],
#                 nodesize     = hyper_grid$nodesize[j],
#                 replace      = hyper_grid$replace[j],
#                 sampsize     = hyper_grid$sampsize[j],
#                 seed         = 123
#               )
# 
# 
#         tf = Sys.time()
# 
#         prediction = predict (fit, newdata = calibration_subset_eval %>% select (- e.cum))
#         truth = calibration_subset_eval$e.cum
# 
#         hyper_grid$prediction[j] = prediction
#         hyper_grid$truth[j] = truth
#         hyper_grid$executation_time[j] = difftime (tf, t0)
# 
#     }
#     
#     df_res = hyper_grid
# 
#     save (df_res, file = paste ("random_forest_cross_validation_results/df_res", i, "Rdata", sep = "."))
# 
# }
```

```{r}
files <- dir ("random_forest_cross_validation_results", pattern = ".Rdata", full.names = TRUE)

res <- do.call (rbind, lapply(files, function(x) get(load(x))))

res %>%

    summarise (rmse = sqrt (mean ((prediction - truth) ^ 2)), 
               .by = c (mtry, nodesize, replace, samp.fraction)) %>%

    arrange (rmse) %>%

    head (n = 20) 
```