### 2024-04-20 scripts/xgboost_cross_validation.qmd

---
title: "xgboost cross-validation"
format:
  html:
    toc: true
    embed-resources: true
---

```{r setup, include=FALSE}
# Options pour le rmarkdown
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

```{r}
library (dplyr)
library (xgboost)
```

```{r}
load (file = "processed_data/data_xgboost.Rdata")
```

```{r}
calibration_subset = data_xgboost %>%
    filter (dataset == "Calibration subset") %>%
    select (- e.rel, - dataset, - pmid)
```

```{r}
# set.seed (123)

# for (i in c (1 : 490)){
# 
#     calibration_subset_train = calibration_subset[- i, ]
# 
#     calibration_subset_eval = calibration_subset[i, ]
# 
#     # create hyper_grid search ------
#     hyper_grid <- expand.grid (
# 
#     max.depth = c (4, 6, 10, 14),
#     nrounds = c (50, 100, 150, 300),
#     eta = c (0.1, 0.2, 0.3),
#     min_child_weight = c (0.5, 1, 2),
#     subsample = c (0.8, 1),
#     prediction = NA,
#     truth = NA,
#     executation_time = NA,
#     row_index = i
#     )
#     # ------
# 
#     for (j in seq_len(nrow(hyper_grid))) {
# 
#         t0 = Sys.time()
# 
#         # fit model for ith hyperparameter combination
#         fit <- xgboost (
# 
#             data = xgb.DMatrix (
# 
#                 data = calibration_subset_train %>%
#                     select (- e.cum) %>%
#                     as.matrix %>%
#                     {.},
# 
#                 label = calibration_subset_train %>%
#                     select (e.cum) %>%
#                     as.matrix %>%
#                     {.}
# 
#             ),
# 
#             max.depth = hyper_grid$max.depth[j],
# 
#             nrounds = hyper_grid$nrounds[j],
# 
#             min_child_weight = hyper_grid$min_child_weight[j],
# 
#             eta = hyper_grid$eta[j],
# 
#             subsample = hyper_grid$subsample[j],
# 
#             verbose = FALSE
# 
#         )
# 
#         tf = Sys.time()
# 
#         prediction = predict (fit, newdata = calibration_subset_eval %>% select (- e.cum) %>% as.matrix)
#         truth = calibration_subset_eval$e.cum
# 
#         hyper_grid$prediction[j] = prediction
#         hyper_grid$truth[j] = truth
#         hyper_grid$executation_time[j] = difftime (tf, t0)
# 
#     }
# 
#     df_res = hyper_grid
# 
#     save (df_res, file = paste ("xgboost_cross_validation_results/df_res", i, "Rdata", sep = "."))
# 
# }
```

```{r}
files <- dir ("xgboost_cross_validation_results", pattern = ".Rdata", full.names = TRUE)

res <- do.call (rbind, lapply(files, function(x) get(load(x))))

res %>%

    summarise (rmse = sqrt (mean ((prediction - truth) ^ 2)), 
               .by = c (max.depth, nrounds, eta, min_child_weight, subsample)) %>%

    arrange (rmse)  %>%

    head (n = 20) 
```