### 2024-05-22 scripts/data_processing.qmd
adding the country variable to the dataset

---
title: "Data processing"
format:
  html:
    toc: true
    embed-resources: true
---

```{r setup, include=FALSE}
# Options pour le rmarkdown
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

In this file, we build several dataframes that are used in the manuscript.qmd file. The following figure summarizes how these dataframes are built.

![](figures/code_explanations.png){width=500, fig-align="center"}

From the files of the ALFAM2 database (available in a Zenodo repository at https://zenodo.org/records/8109634), we first construct data_alfam2, which corresponds to the dataset used in Hafner et al (2019). From data_alfam2, we construct data_ml by retaining only the last observation of each PMID. The dynamics of climatic variables are preserved through class variables. From data_ml, we create data_lasso, data_xgboost, and data_random_forest, all of which contain the same information but are formatted according to the packages used to build the respective models. Additionally, from data_ml, we generate data_scenarios, which correspond to various scenarios on which we test the models. These scenarios are based on quantiles derived from data_ml. From data_scenarios we create four dataframes containing the same information but formatted according to the packages used for the different models.

```{r}
library ("ggplot2") 
library ("dplyr")       
library ("paletteer")       
library ("tidyr")
```

```{r, include = FALSE}
axis_text_size = 24 
axis_title_size = 30
title_size = 26

theme_replace(
    axis.text = element_text(size = axis_text_size), 
    axis.title.x = element_text(size = axis_title_size, angle = 0, margin = ggplot2::margin (t = 10)), 
    axis.title.y = element_text(size = axis_title_size, angle = 90, margin = ggplot2::margin (r = 10)), 
    plot.title = element_text(size = title_size, hjust = 0.5, face = "bold"), 
    legend.text = element_text(size = axis_title_size), 
    legend.title = element_text(size = axis_title_size, face = "bold"), 
    strip.text = element_text(size = axis_title_size, face = "bold", color = "white"), 
    strip.background = element_rect(fill = "#000080"), 
    panel.background = element_rect(fill = "#f3faff"), panel.grid.major = element_line(colour = "white"), 
    panel.spacing = unit(2, "lines")
)

options(ggplot2.discrete.fill = list("#67c5a7"))

Dark2 = paletteer_d("RColorBrewer::Dark2")
```

# data_alfam2

```{r}
# these two files come from the ALFAM2 Zenodo repository (https://zenodo.org/records/8109634)
data_interval_level = read.csv (file = "../data/ALFAM2_interval.csv")
data_plot_level = read.csv (file = "../data/ALFAM2_plot.csv")

# these two files come from the supplementary material of Hafner et al (2019)
calibration_subset_pmids = read.csv (file = "../data/1-s2.0-S1352231018308069-mmc1.csv")
evaluation_subset_pmids = read.csv (file = "../data/1-s2.0-S1352231018308069-mmc2.csv")

pmids = rbind (evaluation_subset_pmids %>% mutate (dataset = "Evaluation subset"),
               calibration_subset_pmids %>% mutate (dataset = "Calibration subset"))
```

```{r}
data_alfam2 = data_interval_level %>%

    filter (pmid %in% pmids$pmid) %>%
    left_join (pmids, by = "pmid") %>%

    filter (ct <= 78) %>%
    select (j.NH3, e.cum, e.rel, pmid, ct, dt, air.temp, wind.2m, rain.rate, rain.cum, dataset)  %>%

    # adding plot level informations --------
    left_join (

        data_plot_level %>%
            filter (pmid %in% pmids$pmid) %>%
            select (pmid, country, tan.app, app.method, app.rate, man.dm, man.ph, man.source, incorp, time.incorp) %>%
            rename (app.mthd = app.method, t.incorp = time.incorp),

        by = "pmid"

    ) %>% 
    # --------

    replace_na (list (rain.rate = 0)) %>% 
    replace_na (list (rain.cum = 0)) %>%

    mutate (wind.2m = replace_na (wind.2m, mean (wind.2m, na.rm = TRUE)), .by = pmid) %>%
    mutate (air.temp = replace_na (air.temp, mean (air.temp, na.rm = TRUE)), .by = pmid) %>%

    rename (time = ct)

save (data_alfam2, file = "processed_data/data_alfam2.Rdata")
```

```{r}
df_join = data_alfam2 %>% select (pmid, country) %>% distinct
```

```{r}
data_alfam2 %>%
  summarise (n_obs = n(), n_trial = length (unique (pmid)))
```

# data_ml

```{r}
data_ml = data_alfam2 %>%

    mutate_if (is.character, as.factor) %>%

    mutate (t.incorp = ifelse (t.incorp <= 0.25, 0, t.incorp)) %>%

    mutate (t.incorp = ifelse (t.incorp == 0 & incorp == "none", 1000, t.incorp)) %>%
    mutate (t.incorp = replace_na (t.incorp, 1000)) %>%

    mutate (man.ph = replace_na (man.ph, median (man.ph, na.rm = TRUE)), .by = man.source) %>%

    # we round dt and time in order to compute the new climatic variables
    # cf the next chunk
    mutate (dt = round (dt, digits = 2)) %>%
    mutate (time = cumsum (dt), .by = pmid) %>%

    {.}
```

```{r}
# data = data_ml
# pmids = data %>% pull (pmid) %>% unique
# 
# data_ml = NULL
# 
# for (i in pmids) {
#         
#     data_tmp = data %>% filter (pmid == i)
#     
#     time_max = max (data_tmp$time)
#     
#     time = seq (0, (time_max - 0.01), by = 0.01)
#     
#     group = c(
#         sapply (c (1 : 5), function (j) rep (j, 400)) %>% as.vector,
#         rep (6, (length (time) - 2000))
#     )
#     
#     # here we added 1+e-8 for numerical reasons, ('19.8 - 16.7 <= 3.1' return FALSE)
#     temp = sapply (time, function (j) data_tmp %>% filter ((time - dt) <= j + 1e-8 & j + 1e-8 < time) %>% pull (air.temp))
#     wind = sapply (time, function (j) data_tmp %>% filter ((time - dt) <= j + 1e-8 & j + 1e-8 < time) %>% pull (wind.2m))
#     rain = sapply (time, function (j) data_tmp %>% filter ((time - dt) <= j + 1e-8 & j + 1e-8 < time) %>% pull (rain.rate))
# 
# 
#     df_temp = data.frame (time = time, temp = temp, group = group) %>% 
# 
#         summarise (temp = mean (temp), .by = group) %>% 
#         pivot_wider (names_from = group, values_from = temp, names_prefix = "temp_")
#                    
# 
#     df_wind = data.frame (time = time, wind = wind, group = group) %>% 
# 
#         summarise (wind = mean (wind), .by = group) %>% 
#         pivot_wider (names_from = group, values_from = wind, names_prefix = "wind_")
#                    
# 
#     df_rain = data.frame (time = time, rain = rain, group = group) %>% 
# 
#         summarise (rain = mean (rain), .by = group) %>% 
#         pivot_wider (names_from = group, values_from = rain, names_prefix = "rain_")
# 
# 
#     data_tmp = cbind (
#         data_tmp[nrow (data_tmp), ] %>% select (dataset, e.cum, e.rel, pmid, country, time, tan.app, app.mthd, app.rate, man.dm, man.ph, man.source, incorp, t.incorp),
#         df_temp,
#         df_wind,
#         df_rain
#     )
#                    
#     data_ml = rbind (data_ml, data_tmp)
#                                                          
# }

# save (data_ml, file = "processed_data/data_ml.Rdata")

load (file = "processed_data/data_ml.Rdata")
```

```{r}
dim (data_ml)
```

## data_random_forest

```{r}
data_random_forest = data_ml %>%

    mutate (man.source = as.numeric (recode (man.source, "pig" = 1, "cat" = 0))) %>%
    mutate (app.mthd = as.numeric (recode (app.mthd, "bc" = 1, "ts" = 2, "os" = 3, "bsth" = 4))) %>%
    mutate (incorp = as.numeric (recode (incorp, "none" = 1, "shallow" = 2, "deep" = 3)))

save (data_random_forest, file = "processed_data/data_random_forest.Rdata")
```

## data_xgboost

```{r}
data_xgboost = data_ml %>%

    mutate (man.source = as.numeric (recode (man.source, "pig" = 1, "cat" = 0))) %>%
    mutate (app.mthd = as.numeric (recode (app.mthd, "bc" = 1, "ts" = 2, "os" = 3, "bsth" = 4))) %>%
    mutate (incorp = as.numeric (recode (incorp, "none" = 1, "shallow" = 2, "deep" = 3)))

save (data_xgboost, file = "processed_data/data_xgboost.Rdata")
```

## data_lasso

```{r}
data_lasso = data_ml %>%

    # making the dummy variables --------
    mutate (man.source.pig = ifelse (man.source == "pig", 1, 0)) %>%
    select (- man.source) %>%

    mutate (app.mthd.bc = ifelse (app.mthd == "bc", 1, 0)) %>%
    mutate (app.mthd.ts = ifelse (app.mthd == "ts", 1, 0)) %>%
    mutate (app.mthd.os = ifelse (app.mthd == "os", 1, 0)) %>%
    select (- app.mthd) %>%

    mutate (incorp.shallow = ifelse (incorp == "shallow", 1, 0)) %>%
    mutate (incorp.deep = ifelse (incorp == "deep", 1, 0)) %>%
    select (- incorp) %>%
    # ---------

    mutate (e.cum = log (e.cum)) %>%

    {.}

save (data_lasso, file = "processed_data/data_lasso.Rdata")
```

# data_scenarios

```{r, fig.width = 10, fig.height = 5}
data_ml %>%

    select (rain_1, rain_2, rain_3, rain_4, rain_5, rain_6) %>%

    apply (2, quantile, probs = seq (0.01, 1, by = 0.01)) %>%
    as.data.frame %>%
    mutate (quantile = row_number () / 100) %>%
    pivot_longer (cols = - c (quantile), names_to = "variable") %>%

    filter (quantile > 0.75) %>%
    ggplot () +
        geom_line (aes (x = quantile, y = value, color = variable)) +
        scale_color_manual (values = Dark2)

```

```{r}
quartiles_climatic_variable = cbind (

    data_ml %>%

    select (temp_1, temp_2, temp_3, temp_4, temp_5, temp_6,
            wind_1, wind_2, wind_3, wind_4, wind_5, wind_6) %>%

    apply (2, quantile, probs = c (0.25, 0.75)) %>%
    as.data.frame,

    data_ml %>%

    select (rain_1, rain_2, rain_3, rain_4, rain_5, rain_6) %>%

    apply (2, quantile, probs = c (0.25, 0.95)) %>%
    as.data.frame
)
```

```{r}
quartiles_climatic_variable
```

```{r}
grid_climatic_variable = quartiles_climatic_variable [, c(1 : 6)] %>% 
    cross_join (quartiles_climatic_variable [, c(7 : 12)]) %>% 
    cross_join (quartiles_climatic_variable [, c(13 : 18)]) %>%

    mutate (group_temp = ifelse (temp_1 == quartiles_climatic_variable[1, 1], "q1", "q2")) %>%
    mutate (group_wind = ifelse (wind_1 == quartiles_climatic_variable[1, 7], "q1", "q2")) %>%
    mutate (group_rain = ifelse (rain_2 == quartiles_climatic_variable[1, 13], "q1", "q2"))
```

```{r}
quartiles_plot_level_variable = data_ml %>%

    select (tan.app, app.rate, man.dm) %>%

    apply (2, quantile, probs = c (0.25, 0.75)) %>%
    as.data.frame
```

```{r}
quartiles_plot_level_variable
```

```{r}
grid_plot_level_variable = expand_grid (
    time = 72,
    tan.app = quartiles_plot_level_variable$tan.app,
    app.rate = quartiles_plot_level_variable$app.rate,
    man.dm = quartiles_plot_level_variable$man.dm,
    man.ph = 7.5,
    man.source = c ("pig", "cat"),
    incorp = c("none", "shallow"),
    app.mthd = c ("bc", "bsth", "os", "ts"),
    t.incorp = c (0, 1000)

) %>%

    filter (! (incorp == "none" & t.incorp %in% c(0))) %>%
    filter (! (incorp == "shallow" & t.incorp == 1000)) %>%
    filter (! (app.mthd %in% c ("bsth", "os", "ts") & incorp == "shallow"))
```

```{r}
nrow (grid_plot_level_variable)
# in the following calcul, the terms in parenthesis correspond to the variables time, tan.app, app.rate, man.dm, man.ph and man.source
# the last terms correspond to the variables incorp, app.mthd, and t.incorp

# incorp = none
(1 * 2 * 2 * 2 * 1 * 2) * 1 * 4 * 1 +

# incorp = shallow
(1 * 2 * 2 * 2 * 1 * 2) * 1 * 1 * 1
```

```{r}
data_scenarios = cross_join (grid_climatic_variable, grid_plot_level_variable) %>%

    mutate (pmid = row_number ()) %>%

    # reordering the variables
    select (pmid, time, tan.app, app.mthd, app.rate, man.dm, man.ph, man.source, incorp, t.incorp,
            temp_1, temp_2, temp_3, temp_4, temp_5, temp_6,
            wind_1, wind_2, wind_3, wind_4, wind_5, wind_6,
            rain_1, rain_2, rain_3, rain_4, rain_5, rain_6,
            group_temp, group_wind, group_rain)

save (data_scenarios, file = "processed_data/data_scenarios.Rdata")
```

## data_scenarios_random_forest
```{r}
data_scenarios_random_forest = data_scenarios %>% 
    select (- pmid, - group_temp, - group_wind, - group_rain) %>%

    mutate (man.source = as.numeric (recode (man.source, "pig" = 1, "cat" = 0))) %>%
    mutate (app.mthd = as.numeric (recode (app.mthd, "bc" = 1, "ts" = 2, "os" = 3, "bsth" = 4))) %>%
    mutate (incorp = as.numeric (recode (incorp, "none" = 1, "shallow" = 2, "deep" = 3))) %>%

    {.}

save (data_scenarios_random_forest, file = "processed_data/data_scenarios_random_forest.Rdata")
```

## data_scenarios_xgboost

```{r}
data_scenarios_xgboost = data_scenarios %>%

    select (- pmid, - group_temp, - group_wind, - group_rain) %>%

    mutate (man.source = as.numeric (recode (man.source, "pig" = 1, "cat" = 0))) %>%
    mutate (app.mthd = as.numeric (recode (app.mthd, "bc" = 1, "ts" = 2, "os" = 3, "bsth" = 4))) %>%
    mutate (incorp = as.numeric (recode (incorp, "none" = 1, "shallow" = 2, "deep" = 3))) %>%

    {.}

save (data_scenarios_xgboost, file = "processed_data/data_scenarios_xgboost.Rdata")
```

## data_scenarios_lasso

```{r}
data_scenarios_lasso = data_scenarios %>%

    select (- pmid, - group_temp, - group_wind, - group_rain) %>%

    # making the dummy variables --------
    mutate (man.source.pig = ifelse (man.source == "pig", 1, 0)) %>%
    select (- man.source) %>%

    mutate (app.mthd.bc = ifelse (app.mthd == "bc", 1, 0)) %>%
    mutate (app.mthd.ts = ifelse (app.mthd == "ts", 1, 0)) %>%
    mutate (app.mthd.os = ifelse (app.mthd == "os", 1, 0)) %>%
    select (- app.mthd) %>%

    mutate (incorp.shallow = ifelse (incorp == "shallow", 1, 0)) %>%
    mutate (incorp.deep = ifelse (incorp == "deep", 1, 0)) %>%
    select (- incorp)
    # ---------

save (data_scenarios_lasso, file = "processed_data/data_scenarios_lasso.Rdata")
```

## data_scenarios_alfam2

```{r}
data_scenarios_alfam2 = data_scenarios %>%

    mutate (t.incorp = na_if (t.incorp, 1000)) %>%

    select (- time) %>%

    pivot_longer (cols = c (10 : 27), names_to = c (".value", "time"), names_pattern = "(.+)_(.+)") %>%

    select (- group_temp, - group_wind, - group_rain) %>%

    rename (air.temp = temp, wind.2m = wind, rain.rate = rain) %>%

    mutate (time = as.numeric (recode (time, "1" = "4", "2" = "8", "3" = "12", "4" = "16", "5" = "20", "6" = "72"))) %>%
    mutate (dt = ifelse (time == 72, (72 - 20), 4), .after = time) %>%

    mutate (rain.cum = cumsum (rain.rate), .by = pmid)

save (data_scenarios_alfam2, file = "processed_data/data_scenarios_alfam2.Rdata")
```